---
title: "Synthetic Experiment. I"
author: "Roberto Maestre"
date: "12/17/2018"
output: github_document
---

```{r setup, include=FALSE}
library(variableStars)
library(data.table)
library(ggplot2)
library(ggsci)
library(microbenchmark)
library(RColorBrewer)
```


## Experiment configuration

#### Parameters for execution

```{r parameters}
paramters = list(
  "filter" = "uniform",
  "gRegimen" = 0,
  "minDnu" = 15,
  "maxDnu" = 95,
  "dnuValue" = -1,
  "dnuGuessError" = 10,
  "dnuEstimation" = TRUE,
  "numFrequencies" = 30,
  "debug" = TRUE)
```

#### Data source

Choose: 

 * dataFlag = T, to use synthetic data generated or,

 * dataFlag = F, to use a real pulsar photometry

```{r flag}
dataFlag = F
```


#### Frequences and Amplitudes on photometry data

```{r calculateEspectrum, warning=F}
# Generate first pattern
dt.spectrum <- data.frame(
  "frequency" = seq(from=0, to=10, by=0.5),
  "amplitude" = 10
)
# Generate second pattern as the biased first
dt.spectrum.bias <- data.frame(dt.spectrum)
dt.spectrum.bias$frequency <- dt.spectrum.bias$frequency + 0.3
dt.spectrum.bias$amplitude <- 5

# All together
dt.spectrum <- rbind(dt.spectrum, dt.spectrum.bias)

# Get max amplitude
maxAmplitude <- dt.spectrum[which.max(dt.spectrum$amplitude), ]
# Plot amplitudes
plot_spectrum(min(dt.spectrum$frequency),
              max(dt.spectrum$frequency),
              dt.spectrum)
# Save Data to disk (to be replicated)
write.table(
  dt.spectrum[c("frequency", "amplitude")],
  file = "/tmp/data.csv",
  sep = "\t",
  quote = F,
  row.names = F,
  col.names = F
)
```

##  Experiment execution

process is the main method on the variableStars package to compute and estimate all parameters

```{r calculate, warning=F}
result <- process(
  dt.spectrum$frequency,
  dt.spectrum$amplitude,
  filter = "uniform",
  gRegimen = 0,
  minDnu = 15,
  maxDnu = 95,
  dnuValue = -1,
  dnuGuessError = 10,
  dnuEstimation = TRUE,
  numFrequencies = 30,
  debug = TRUE
)
```

#### Main results

* **Dnu** (```result$dnu```) = `r result$dnu`.
* **DnuGuess** ```(result$dnuGuess```) = `r result$dnuGuess`.
* **DnuPeak** (```result$dnuPeak```) = `r result$dnuPeak`.
* **Frequency** (```result$photometry$frequency```) = `r head(c(result$photometry$frequency), 25)`...
* **Amplitude** (```result$photometry$amplitude```) = `r head(c(result$photometry$amplitude), 25)`...
* **Diffs** (```result$diffHistogram$diffs```) = `r head(c(result$diffHistogram$diffs), 25)`...

#### Apodization


```{r apodization, warning=F}
# Plot frecuency and amplitude
ggplot(
  aes(x = frequences, y = amplitude),
  data = data.frame(
    "frequences" = result$apodization$frequences,
    "amplitude" = result$apodization$amp
  )
) +
  geom_point() +
  geom_line() +
  ggtitle("Apodization- Frequences and amplitudes") +
  theme_bw()
```

#### Periodicities

```{r ftPower, warning=F}
# DS to save all data
dt <- setNames(data.frame(matrix(ncol = 3, nrow = 0)), c("fInv", "b", "label"))
# Prepare data
ranges <- names(result$fresAmps)
for(range in ranges){
  dt <- rbind(dt, 
  data.frame("fInv"=result$fresAmps[[as.character(range)]][["fInv"]],
             "b"=result$fresAmps[[as.character(range)]][["b"]],
            "label"=paste(result$fresAmps[[as.character(range)]][["label"]]," freqs")))
}
# Plot frecuency and amplitude
ggplot(aes(x = fInv, y = b, group=label, colour=label), data = dt) +
  #geom_point(alpha=0.2) +
  geom_line(alpha=0.8) +
  ggtitle(expression(paste("Periodicities (",d^-1,")"))) +
  xlab(expression(paste("Periodicities (",mu,"hz)"))) +
  ylab("Amplitude") +
  theme_bw() + 
  scale_color_lancet() +
  xlim(0, 10)
```

####  Histogram fo differences. 

We only show bins with >0 values


```{r diffsHistogram, warning=F}
dt <- data.frame(result$diffHistogram$histogram)
ggplot(aes(x = bins, y = values), data = dt[dt$values > 0,]) +
  geom_bar(stat = "identity") +
  ggtitle("Histogram of differences") +
  theme_bw()
```


### Autocorrelation

```{r autocor}
dt <- data.frame(result$crossCorrelation)

ggplot(aes(x = index, y = autocorre), data = dt) +
  geom_line(stat = "identity") +
  ggtitle("Autocorrelacion (Crosscorrelation)") +
  xlab(expression(paste("Periodicities (",mu,"hz)"))) +
  ylab("Autocorrelation") +
  theme_bw()
```

#### For first all frecuencies


```{r echelle30, warning=F, message=F, cache=F}
dt <- data.frame(
  "x" = result$echelle$modDnuStacked,
  "y" = result$echelle$freMas,
  "h" = result$echelle$amplitudes
)
plot_echelle(dt)
```

